---
title: "Bioregionalization of the area selected for the comparsion study using the Infomap algorithm"
output: html_notebook
---

<!-- Person -->
```{r}
# person  <- 'Genoveva'
person  <- 'Ruben'

if (person == 'Genoveva') {source('sc_func_Infomap.R')}
if (person == 'Ruben') {source('01_Analyses/sc_func_Infomap.R')}
```

# Prep
```{r}
library(tidyverse)
library(sp)
```

# Data
```{r}
da <- otu_new_cs
names(da)<- c("SampID", "clean_taxonomy",  "count","density_n100m2", "VL")
da <- data.frame(da)
head(da)
```
# Data summary
```{r}
df <- data.frame(Samples = length(unique(da$SampID)),
                 Species = length(unique(da$clean_taxonomy)),
                  Video_lines = length(unique(da$VL)))
df
```
# Write input infomap  
```{r}
path_write_Input <-  "regionalization"
Input_name <- 'Ip_Benthos.net'
edges_file <- 'edge_Benthos'
# Create input in the same folder
write.input.info(da, path_write_Input, Input_name, edges_file, weighted = T)
```
# Run Infomap
```{r}
path.info <- "infomap-ubuntu"

# Directorio donde esta el input
path.in <- "regionalization"

# Directorio donde quieres guardar el output
path.out <- "regionalization"
          
# -N X = numero de analisis
# -s = seed
# empty = hierarchical // -2 = two-levels
N_trials <- 100 # Yo suelo poner 100
seed <- 1
# Si quieres hacer una biorregionalizacion a un solo nivel o explorar la existencia de multiples niveles jerarquicos
Hierarchical <- F
if (Hierarchical == T){
  Type <- ''
} 
if (Hierarchical == F){
  Type <- '-2'
} 
          
parameters <- paste("-N", N_trials, "-s", 1, Type)
      
# Run
par.n <- run.infomap(path.info, path.in, Input_name, path.out, parameters)
         
```

# Explore saved output
```{r}
dir.info.out <- 'regionalization/Ip_Benthos.tree'
dir.edges <- 'regionalization/edge_Benthos'
            
# :::::::::::::: 
# Only the table
# :::::::::::::: 
db_output <- info.out(dir.info.out)
db_output
        
# ::::::::::::::::::::::::::::::::
# Table plus metrics for the roles
# ::::::::::::::::::::::::::::::::
Output <- func_Link_Info(dir.info.out, dir.edges, Hierarchical_level = 'last_mod', show.add.metrics = T, ret='all', links_as_cell_spp = T)
# Esto genera un output de 4 archivos
names(Output)
# Output of Infomap
info <- Output[[which(names(Output)=='Infomap')]]
# Links and associated modules
db_Links <- Output[[which(names(Output)=='links')]]
# Indicative values of the species of each Module
db_IndVal <- Output[[which(names(Output)=='db_IndVal')]]
# Metrics to estimate the biogeographical roles
db_Metrics <- Output[[which(names(Output)=='db_Metrics')]]
db_Metrics <- db_Metrics %>%  
  dplyr::select(cell, Module, Rel_Species_Richness, Biota_Overlap, Endemicity, Rel_Occupancy)

db_Metrics
```
# Loop bioregionalization to look at stability (aka solution landscape)
```{r}
all_classif <- data.frame(cell=unique(da$SampID))


for (i in 1:20) {
  # Create a new column name dynamically
  new_col_name <- paste("Result_Run", i, sep = "_")
  
  # Apply the function to the "Value" column and store the result in the new column
  parameters <- paste("-N", N_trials, i, 1, Type)
  run.infomap(path.info, path.in, Input_name, path.out, parameters)
  db_output <- info.out(dir.info.out)
  db_Metrics <- Output[[which(names(Output)=='db_Metrics')]]
  all_classif[[new_col_name]] <- db_Metrics %>% dplyr::select(Module)
}

```



# Biogeographical sectors
```{r}
m<- db_Metrics %>% filter(complete.cases(.)) %>% dplyr::select(Rel_Species_Richness, Biota_Overlap, Endemicity, Rel_Occupancy) %>%
   as.matrix


cl <- kmeans(m,7)

idx <- which(!(complete.cases(db_Metrics)))

insertValuesAtIdx <- function(vec,values,idx)
{
  res<-vector(mode=mode(vec),length = length(vec)+length(idx))
  res[-idx]<-vec
  res[idx]<-values
  return(res)
}

roles <- insertValuesAtIdx(cl$cluster, NA, idx)

db_Metrics <- db_Metrics %>% mutate(role=roles)
db_Metrics

```

# Export and visualize
```{r}
library(sf)

db_Spatial <- db_Metrics %>% mutate(x_coord = cell %l% data.frame(samp_info_cs$SampID2, samp_info_cs$x_coord)) %>%
  mutate(y_coord = cell %l% data.frame(samp_info_cs$SampID2, samp_info_cs$y_coord)) %>%
  filter(!is.na(x_coord))

db_Spatial <- db_Spatial %>% st_as_sf(coords = c("x_coord","y_coord"))%>% st_set_crs("EPSG:32633")
st_write(db_Spatial, ("infomap_result_new2.shp"), append=FALSE)
        
sort(table(db_Spatial$Module))

```
# Average distance to point of different class, by class
```{r}
db_Spatial_sp <- as_Spatial(db_Spatial)

dist_matrix <- spDists(db_Spatial_sp, longlat = FALSE)

# Create a dataframe to store the average distance for each point
db_Spatial_sp$avg_dist_same_class <- NA

# Loop through each point
for (i in 1:nrow(db_Spatial_sp)) {
  # Find indices of points with the same class as the current point
  same_class_indices <- which(db_Spatial_sp$Module == db_Spatial_sp$Module[i])
  
  # Exclude the point itself (distance 0)
  same_class_indices <- same_class_indices[same_class_indices != i]
  
  # Calculate the average distance to points of the same class
  if (length(same_class_indices) > 0) {
    db_Spatial_sp$avg_dist_same_class[i] <- mean(dist_matrix[i, same_class_indices])
  } else {
    db_Spatial_sp$avg_dist_same_class[i] <- NA  # If there are no other points of the same class
  }
}

# Calculate the average distance per class
avg_dist_per_class <- aggregate(avg_dist_same_class ~ Module, data = db_Spatial_sp, FUN = mean, na.rm = TRUE)

# View the results
arrange(avg_dist_per_class, avg_dist_same_class)
```
# Variograms
```{r}
class=29

# Load required libraries
library(gstat)
library(sp)


# Create indicator variables for the categories
db_Spatial_sp$A <- as.numeric(db_Spatial_sp$Module == class)
db_Spatial_sp$B <- as.numeric(db_Spatial_sp$Module != class)

# Calculate the variogram for the "A" category
v_A <- variogram(A ~ 1, db_Spatial_sp)

# Calculate the variogram for the "B" category
v_B <- variogram(B ~ 1, db_Spatial_sp)

# Plot the variograms
plot(v_A, main = "Indicator Variogram for Category A")
#plot(v_B, main = "Indicator Variogram for Category B")

```

# Explore classes
```{r}
class=2
db_Metrics[which(db_Metrics$Module==class),]
```
# Explore classes
```{r}
class=29
db_output[which(db_output$Module_1==class),]
```
# Species?
```{r}
p="1749_01"
da[which(da$SampID==p),]
```

# Module assignation based on codelength
```{r}
#input: new localities with their species based on 
```

# Are modules different in their env profiles?
```{r}
#multinomial regression 
```

# Multilayer regionalization
```{r}
#1 capa: datos bioticos
#2 capa: datos abioticos

#output: tabla localidades - module
```

# climatic bins (no tiene en cuenta las especies)
```{r}
#crear climatic bins
#calcular localidad - bin climatico
#output: table nodos - modules
#(elife paper)

```

